**Sequencing 16S rRNA gene fragments using the PacBio SMRT DNA sequencing system**

### 2016-01-27 Schloss_PacBio16S_PeerJ_2016.Rmd
Complete code edits for Results section

min.region <- paste(pretty.region[names(rnd.reduction[rnd.reduction==min.rnd.red])],collapse=" and ")
max.region <- pretty.region[[names(rnd.reduction[rnd.reduction==max.rnd.red])]]
```

***Pre-clustering sequences to further reduce sequencing noise.***
Previously, we implemented a pre-clustering algorithm where sequences were sorted
by their abundance in decreasing order and rare sequences are clustered with a
more abundant sequence if the rare sequences have fewer mismatches than a
defined threshold when compared to the more abundant sequence [refs]. The recommended
threshold was a 1-nt difference per 100-nt of sequence data. For example, the
threshold for 250 bp fragment from the V4 region would be 2 nt or 14 for the
1458 bp V1-V9 fragments. This approach removes residual PCR and sequencing
errors while not overwhelming the resolution needed to identify OTUs that are
based on a 3% distance threshold. The tradeoff of this approach is that one would
unable to differentiate V1-V9 sequences that truly differed by less than 14 nt.
When we applied this approach to our PacBio data, we observed a reduction in the error rate
between `r min.rnd.red` (`r min.region`) and `r max.rnd.red`% (`r max.region`).
The final error rates varied between
`r round(100*min(pooled_error[,"precluster"]), digits=2)` (`r names(which.min(pooled_error[,"precluster"]))`) and `r round(100*max(pooled_error[,"precluster"]), digits=2)`% (`r names(which.max(pooled_error[,"precluster"]))`).
 The full-length (i.e. V1V9) fragments had an error rate of
`r round(100* pooled_error["V1V9","precluster"], digits=2)`%
(Figure 2B), this is similar to what we have previously observed using the 454 and
Illumina MiSeq platforms (0.02%){Schloss, 2011 #2466; Kozich, 2013 #2719}

```{r}
getOTUResults <- function(region){
  perfect <- read.table(file=paste0("analysis/", region, "/HMP_MOCK.filter.pick.phylip.an.summary"), header=T)
  nochims <- read.table(file=paste0("analysis/", region, "/", region, ".mock.precluster.perfect.pick.an.ave-std.summary"), header=T)
  observed <- read.table(file=paste0("analysis/",region, "/", region, ".trim.unique.good.filter.unique.precluster.pick.an.merge.groups.ave-std.summary"), header=T)
	obs.mock <- observed[observed$group=="mock"   & observed$method=="ave", "sobs"]
	obs.soil <- observed[observed$group=="soil"   & observed$method=="ave", "sobs"]
	obs.mouse <- observed[observed$group=="mouse" & observed$method=="ave", "sobs"]
	obs.human <- observed[observed$group=="human" & observed$method=="ave", "sobs"]
	nseqs <- nochims[nochims$method=="ave", "nseqs"]

	return(c(perfect[1,2], nochims[1,3], obs.mock, obs.soil, obs.mouse, obs.human, nseqs))
}

otu.table <- t(sapply(regions, getOTUResults))
nseqs <- otu.table[1,7]
perfect.otu <- otu.table[1,1]
otu.table <- otu.table[,-c(1,7)]

otu.table <- cbind(100 * final.error[rownames(otu.table),"precluster.error"], otu.table)
colnames(otu.table) <- c("Error Rate (%)", "Mock with no chimeras", "Observed mock", "Soil", "Mouse", "Human")
rownames(otu.table) <- pretty.region[rownames(otu.table)]

@@

#|      | Error Rate (%)| Mock with no chimeras| Observed mock|  Soil| Mouse| Human|
#|:-----|--------------:|---------------------:|-------------:|-----:|-----:|-----:|
#|V4    |           0.14|                  26.0|          27.4| 335.8|  78.5|  58.1|
#|V1-V3 |           0.77|                  53.6|          77.0| 436.6| 159.7| 116.8|
#|V3-V5 |           0.83|                  83.1|         106.8| 388.9| 259.6|  91.6|
#|V1-V5 |           0.26|                  34.6|          40.7| 403.2|  88.8|  74.7|
#|V1-V6 |           0.57|                  57.4|          89.7| 431.9| 138.3|  95.9|
#|V1-V9 |           0.32|                  29.5|          41.8| 390.9|  81.9| 102.0|

getCoverage <- function(region){
  observed <- read.table(file=paste0("analysis/",region, "/", region, ".trim.unique.good.filter.unique.precluster.pick.an.merge.groups.ave-std.summary"), header=T)
	obs.mock <- observed[observed$group=="mock"   & observed$method=="ave", "coverage"]
	obs.soil <- observed[observed$group=="soil"   & observed$method=="ave", "coverage"]
	obs.mouse <- observed[observed$group=="mouse" & observed$method=="ave", "coverage"]
	obs.human <- observed[observed$group=="human" & observed$method=="ave", "coverage"]

	return(c(obs.mock, obs.soil, obs.mouse, obs.human))
}

coverage.table <- t(sapply(regions, getCoverage))
colnames(coverage.table) <- c("Mock", "Soil", "Mouse", "Human")
rownames(coverage.table) <- pretty.region[rownames(coverage.table)]
```

***Effects of error rates on OTU assignments.***
The sequencing error rate is known to affect the number of OTUs that are
observed {Schloss, 2011 #2466}. For each region, we determined that if there were no chimeras or
PCR or sequencing errors, then we would expect to find `r perfect.otu[[1]]` OTUs. When
achieved perfect chimera removal, but allowed for PCR and sequencing errors, we
observed between `r diff<-otu.table[,2]-perfect.otu;
round(min(diff), digits=1)` (`r names(which.min(diff))`) and `r round(max(diff), digits=1)`
(`r names(which.max(diff))`) extra OTUs. The range in the number of extra OTUs was largely
explained by the sequencing error rate (Pearson's R=`r round(cor.test(diff, final.error[,"precluster.error"])$estimate, 2)`). Next, we determined
the number of OTUs that were observed when we used UCHIME to identify chimeric sequence.
Under these more realistic conditions, we observed between `r diff<-otu.table[,3]-perfect.otu;
round(min(diff), digits=1)` (`r names(which.min(diff))`) and `r round(max(diff), digits=1)`
(`r names(which.max(diff))`) extra OTUs. Finally, we calculated the number of OTUs in the
soil, mouse, and human samples using the same pipeline with chimera detection and removal
based on the UCHIME algorithm. Again, we found that there was a strong correlation between
the number of observed OTUs and the error rate for the
soil (R=`r format(round(cor.test(otu.table[,"Soil"], final.error[,"precluster.error"])$estimate, 2), nsmall=2)`),
mouse (R=`r format(round(cor.test(otu.table[,"Mouse"], final.error[,"precluster.error"])$estimate, 2), nsmall=2)`),
and human samples (R=`r format(round(cor.test(otu.table[,"Human"], final.error[,"precluster.error"])$estimate, 2), nsmall=2)`).
These results underscore the effect of sequencing error on the inflation of the number of
observed OTUs.

```{r, classificationSummary}
data <- read.table(file="taxonomy.depth.analysis", header=T)

rdp <- cbind("Mock"=data[data$database=="rdp" & data$sample=="mock", "total"],
						 "Human"=data[data$database=="rdp" & data$sample=="human", "total"],
						 "Mouse"=data[data$database=="rdp" & data$sample=="mouse", "total"],
						 "Soil"=data[data$database=="rdp" & data$sample=="soil", "total"])
rownames(rdp) <- c("V4", "V3-V5", "V1-V3", "V1-V5", "V1-V6", "V1-V9")

gg <- cbind("Mock"=data[data$database=="gg" & data$sample=="mock", "total"],
						"Human"=data[data$database=="gg" & data$sample=="human", "total"],
						"Mouse"=data[data$database=="gg" & data$sample=="mouse", "total"],
						"Soil"=data[data$database=="gg" & data$sample=="soil", "total"])
rownames(gg) <- c("V4", "V3-V5", "V1-V3", "V1-V5", "V1-V6", "V1-V9")

silva <- cbind("Mock"=data[data$database=="silva" & data$sample=="mock", "total"],
							 "Human"=data[data$database=="silva" & data$sample=="human", "total"],
							 "Mouse"=data[data$database=="silva" & data$sample=="mouse", "total"],
							 "Soil"=data[data$database=="silva" & data$sample=="soil", "total"])
rownames(silva) <- c("V4", "V3-V5", "V1-V3", "V1-V5", "V1-V6", "V1-V9")

gg.sp <- cbind("Mock"=data[data$database=="gg" & data$sample=="mock", "X7"],
							 "Human"=data[data$database=="gg" & data$sample=="human", "X7"],
							 "Mouse"=data[data$database=="gg" & data$sample=="mouse", "X7"],
							 "Soil"=data[data$database=="gg" & data$sample=="soil", "X7"])
rownames(gg.sp) <- c("V4", "V3-V5", "V1-V3", "V1-V5", "V1-V6", "V1-V9")
```

@@

```{r}
error <- read.table(file="analysis/v4/v4.trim.unique.good.filter.unique.pick.error.summary", header=T, row.names=1)
error.nochim <- error[error$numparents==1,]
one.off <- error.nochim[error.nochim$mismatches==1,]
one.off.sum <- sum(one.off$weight)
v4.table <- table(one.off$weight)
v4.total <- sum(error$weight)

error <- read.table(file="analysis/v35/v35.trim.unique.good.filter.unique.pick.error.summary", header=T, row.names=1)
error.nochim <- error[error$numparents==1,]
one.off <- error.nochim[error.nochim$mismatches==1,]
one.off.sum <- sum(one.off$weight)
v35.table <- table(one.off$weight)
v35.total <- sum(error$weight)

error <- read.table(file="analysis/v15/v15.trim.unique.good.filter.unique.pick.error.summary", header=T, row.names=1)
error.nochim <- error[error$numparents==1,]
one.off <- error.nochim[error.nochim$mismatches==1,]
one.off.sum <- sum(one.off$weight)
v15.table <- table(one.off$weight)
v15.total <- sum(error$weight)
```

***Sequencing errors are not random.***
Above, we described that although there was no obvious bias in the substitution
or insertion rate, we did observe that guanines and cytosines were more likely
to be deleted than adenines and thymidines. This lack of randomness in the error
profile suggested that there might be a systematic non-random distribution of
the errors across the sequences. This would manifest itself by the creation of
duplicate sequences with the same error. Because we were able to obtain a large
number of reads from the mock communities where we sequenced the V4
(N=`r v4.total`), V1-V5 (N=`r v15.total` sequences), and V3-V5 (N=`r v35.total`)
regions, we investigated the mock community data from these regions further. We
identified all of the sequences that had a 1-nt difference to the true sequence.
For these three regions, a majority of the sequences with 1-nt errors were only
observed once (V4: `r round(100*v4.table[1]/sum(v4.table), 1)`%, V1-V5:
`r round(100*v15.table[1]/sum(v15.table), 1)`%, V3-V5:
`r round(100*v35.table[1]/sum(v35.table), 1)`%). We found that the frequency of the
most abundant 1-nt error paralleled the number of sequences. There were two sequences
in the V4 dataset that occurred `r names(v4.table[length(v4.table)])` times, one
sequence in the V1-V5 dataset that occurred `r names(v15.table[length(v15.table)])`
times, and one sequence in the V3-V5 dataset that occurred
`r names(v35.table[length(v35.table)])` times. Contrary to previous reports
{Carneiro, 2012 #3458;Koren, 2012 #3459}, these results indicate that
reproducible errors occur with the PacBio sequencing platform and that they can be
quite frequent.

## Conclusions
The various sequencing platforms that are available to microbial ecologists are
able to fill unique needs and have their own strengths and weaknesses. For
sequencing the 16S rRNA gene, the 454 platform is able to generate a moderate
number of high-quality 500-nt sequence fragments (error rates below 0.02%) {Schloss, 2011 #2466} and
the MiSeq platform is able to generate a large number of high-quality 250-nt
sequence fragments (error rates below 0.02%) {Kozich, 2013 #2719}. The promise of the PacBio
sequencing platform was the generation of high-quality near full-length sequence
fragments. As we have shown in this study, it is possible to generate near
full-length sequences; however, the error rate associated with those reads is
considerable (i.e. `r round(100*final.error["v19","precluster.error"],2)`%)
and requires a level of sequencing coverage that is not commonly observed in a
typical sequencing run. This results in the generation of a small number of low
quality full-length sequences. When we considered the shorter V4 region, which is
similar in length to what is sequenced by the MiSeq platform, the error rates
we observed with the PacBio platform were nearly 5-fold higher than what has
previously been reported. It appears that the promise offered by the PacBio
platform has not been realized.

@@

oligos <- read.table(file="pacbio.oligos", stringsAsFactors=F)[1:6,]
rownames(oligos) <- pretty.region[oligos$V4]
oligos <- oligos[,c("V2", "V3")]
colnames(oligos) <- c("Forward", "Reverse")

@@

```{r Table_2, warning=TRUE, results='asis'}
rownames(reasons.lost) <- pretty.region[rownames(reasons.lost)]
reasons.lost <- cbind(orig.nseqs, 100*reasons.lost, basic.nseqs)
knitr::kable(reasons.lost, digits=1, row.names=TRUE,
						 col.names=c("Initial sequences (N)", "Good reads (%)",
						 						"Wrong start/end position (%)",
						 						"Excessively long homopolymers (%)",
						 						"Ambiguous base calls (%)",
						 						"Sequences remaining (N\\)"),
						 align="c")
```

@@

par(mar=c(5,2,0.5, 0.5))
plot(1, xlim=c(2,30), ylim=c(0,0.05), type="n", xlab="", ylab="", axes=F)
axis(2, at=seq(0,0.05, 0.01), label=FALSE, tick=T, las=2)
#axis(2, at=seq(0,0.05, 0.01), label=seq(0,5, 1), las=2)
axis(1, cex.axis=1.5)
lapply(regions, plotLines)
legend(x=18, y=0.05, legend=pretty.region[regions], lty=1, lwd=2, col=clrs[regions])
mtext(side=2, line=0.5, at=0.05, text="C", cex=2, font=2, las=2)
title(xlab="Coverage", ylab="", cex.lab=1.5)
box()

@@

legend(x=10, y=0.50, legend=pretty.region[regions], lty=1, lwd=2, col=clrs[regions], cex=1.0)
```

>>>>

min.region <- pretty_region[names(reduction[which.min(reduction)])]
max.region <- pretty_region[names(reduction[which.max(reduction)])]
```

***Pre-clustering sequences to further reduce sequencing noise.***
Previously, we implemented a pre-clustering algorithm where sequences were sorted
by their abundance in decreasing order and rare sequences are clustered with a
more abundant sequence if the rare sequences have fewer mismatches than a
defined threshold when compared to the more abundant sequence [@refs]. The recommended
threshold was a 1-nt difference per 100-nt of sequence data. For example, the
threshold for 250 bp fragment from the V4 region would be 2 nt or 14 for the
1458 bp V1-V9 fragments. This approach removes residual PCR and sequencing
errors while not overwhelming the resolution needed to identify OTUs that are
based on a 3% distance threshold. The tradeoff of this approach is that one would
unable to differentiate V1-V9 sequences that truly differed by less than 14 nt.
When we applied this approach to our PacBio data, we observed a reduction in the error rate
between `r min(reduction)` (`r min.region`) and `r max(reduction)`% (`r max.region`).
The final error rates varied between
`r round(100*min(pooled_error[,"precluster"]), digits=2)` (`r names(which.min(pooled_error[,"precluster"]))`) and `r round(100*max(pooled_error[,"precluster"]), digits=2)`% (`r names(which.max(pooled_error[,"precluster"]))`).
 The full-length (i.e. V1V9) fragments had an error rate of
`r round(100* pooled_error["V1V9","precluster"], digits=2)`%
(Figure 2B), this is similar to what we have previously observed using the 454 and
Illumina MiSeq platforms (0.02%){Schloss, 2011 #2466; Kozich, 2013 #2719}

```{r}
getAveSobs <- function(region, sample){
	file_name <- paste0("data/mothur_pool/", region, ".", sample, ".screen.unique.good.filter.unique.precluster.pick.an.ave-std.summary")

	data_sobs <- NA

	if(file.exists(file_name)){
		data <- read.table(file=file_name, header=T)
		data_sobs <- data[data$method=="ave", "sobs"]
	}

	data_sobs
}

getOTUResults <- function(region){
 	perfect <- read.table(file=paste0("data/mothur_pool/HMP_MOCK.", region, ".pick.phylip.an.summary"), header=T)
 	nochims <- read.table(file=paste0("data/mothur_pool/", region, ".mock.screen.unique.good.filter.unique.precluster.perfect.an.ave-std.summary"), header=T)

 	obs.mock <- getAveSobs(region, "mock")
 	obs.mouse <- getAveSobs(region, "mouse")
 	obs.human <- getAveSobs(region, "human")
 	obs.soil <- getAveSobs(region, "soil")

	nseqs <- nochims[nochims$method=="ave", "nseqs"]

	return(c(perfect[1,2], nochims[1,"sobs"], obs.mock, obs.soil, obs.mouse, obs.human, nseqs))
}

otu.table <- t(sapply(regions, getOTUResults))
nseqs <- otu.table[[1,7]]
perfect.otu <- otu.table[1,1]
otu.table <- otu.table[,-c(1,7)]

otu.table <- cbind(100 * pooled_error[rownames(otu.table),"precluster"], otu.table)
colnames(otu.table) <- c("Error Rate (%)", "Mock with no chimeras", "Observed mock", "Soil", "Mouse", "Human")
rownames(otu.table) <- pretty_region[rownames(otu.table)]

@@

#|      | Error Rate (%)| Mock with no chimeras| Observed mock|  Soil| Mouse| Human|
#|:-----|--------------:|---------------------:|-------------:|-----:|-----:|-----:|
#|V4    |           0.16|                  33.4|          48.7| 630.1| 212.4| 117.0|
#|V1-V3 |           0.07|                  20.5|          27.7| 726.3| 107.5| 105.4|
#|V3-V5 |           0.04|                  21.1|          31.2|    NA| 112.3|  83.7|
#|V1-V5 |           0.02|                  19.8|          27.2| 694.4|  66.0|  85.3|
#|V1-V6 |           0.03|                  20.0|          46.4| 693.3|  75.5|  89.2|
#|V1-V9 |           0.03|                  19.5|          40.3|    NA| 102.7| 117.6|

```

***Effects of error rates on OTU assignments.***
The sequencing error rate is known to affect the number of OTUs that are
observed {Schloss, 2011 #2466}. For each region, we determined that if there were no chimeras or
PCR or sequencing errors, then we would expect to find `r perfect.otu[[1]]` OTUs. When
achieved perfect chimera removal, but allowed for PCR and sequencing errors, we
observed between `r diff<-otu.table[,2]-perfect.otu;
round(min(diff), digits=1)` (`r names(which.min(diff))`) and `r round(max(diff), digits=1)`
(`r names(which.max(diff))`) extra OTUs. The range in the number of extra OTUs was largely
explained by the sequencing error rate (Pearson's R=`r round(cor.test(diff, pooled_error[,"precluster"])$estimate, 2)`). Next, we determined
the number of OTUs that were observed when we used UCHIME to identify chimeric sequence.
Under these more realistic conditions, we observed between `r diff<-otu.table[,3]-perfect.otu;
round(min(diff), digits=1)` (`r names(which.min(diff))`) and `r round(max(diff), digits=1)`
(`r names(which.max(diff))`) extra OTUs. Finally, we calculated the number of OTUs in the
soil, mouse, and human samples using the same pipeline with chimera detection and removal
based on the UCHIME algorithm. Suprisingly, there was not a clear relationship across sample type and region. Again, we found that there was a strong correlation between
the number of observed OTUs and the error rate for the
mouse (R=`r format(round(cor.test(otu.table[,"Mouse"], pooled_error[,"precluster"])$estimate, 2), nsmall=2)`)
and human samples (R=`r format(round(cor.test(otu.table[,"Human"], pooled_error[,"precluster"])$estimate, 2), nsmall=2)`).
These results underscore the effect of sequencing error on the inflation of the number of
observed OTUs.

```{r, classificationSummary}
data <- read.table(file="data/process/taxonomy.depth.analysis", header=T)

rdp <- 100 * cbind("Mock"=data[data$db=="pds" & data$dataset=="mock", "genus_species"],
				 "Human"=data[data$db=="pds" & data$dataset=="human", "genus_species"],
				 "Mouse"=data[data$db=="pds" & data$dataset=="mouse", "genus_species"],
				 "Soil"=data[data$db=="pds" & data$dataset=="soil", "genus_species"])
rownames(rdp) <- c("V4", "V3-V5", "V1-V3", "V1-V5", "V1-V6", "V1-V9")

gg <- 100 * cbind("Mock"=data[data$db=="gg" & data$dataset=="mock", "genus_species"],
				"Human"=data[data$db=="gg" & data$dataset=="human", "genus_species"],
				"Mouse"=data[data$db=="gg" & data$dataset=="mouse", "genus_species"],
				"Soil"=data[data$db=="gg" & data$dataset=="soil", "genus_species"])
rownames(gg) <- c("V4", "V3-V5", "V1-V3", "V1-V5", "V1-V6", "V1-V9")

silva <- 100 * cbind("Mock"=data[data$db=="nr" & data$dataset=="mock", "genus_species"],
					 "Human"=data[data$db=="nr" & data$dataset=="human", "genus_species"],
					 "Mouse"=data[data$db=="nr" & data$dataset=="mouse", "genus_species"],
					 "Soil"=data[data$db=="nr" & data$dataset=="soil", "genus_species"])
rownames(silva) <- c("V4", "V3-V5", "V1-V3", "V1-V5", "V1-V6", "V1-V9")

gg.sp <- 100 * cbind("Mock"=data[data$db=="gg" & data$dataset=="mock", "X7"],
					 "Human"=data[data$db=="gg" & data$dataset=="human", "X7"],
					 "Mouse"=data[data$db=="gg" & data$dataset=="mouse", "X7"],
					 "Soil"=data[data$db=="gg" & data$dataset=="soil", "X7"])
rownames(gg.sp) <- c("V4", "V3-V5", "V1-V3", "V1-V5", "V1-V6", "V1-V9")
```

@@

```{r}

getRegionNonRandom <- function(region){
	file_name <- paste0("data/mothur_pool/", region, ".mock.screen.unique.good.filter.unique.error.summary")

	error <- read.table(file=file_name, header=T, row.names=1)
	error.nochim <- error[error$numparents==1,]

	one.off <- error.nochim[error.nochim$mismatches==1,]

	one.off.n <- sum(one.off$weight)
	one.off.table <- table(one.off$weight)
	nseqs <- sum(error.nochim$weight)
	one.off.table
}

non_random <- lapply(regions, getRegionNonRandom)
names(non_random) <- regions

percent_one_off <- 100 * sapply(non_random, function(x){x[1]/sum(x)})

```

***Sequencing errors are not random.***
Above, we described that although there was no obvious bias in the substitution
or insertion rate, we did observe that guanines and adenines were more likely to
be deleted than cytosines or thymidines. This lack of randomness in the error
profile suggested that there might be a systematic non-random distribution of
the errors across the sequences. This would manifest itself by the creation of
duplicate sequences with the same error. We identified all of the mock community
sequences that had a 1-nt difference to the true sequence. For these three
regions, between `r min(percent_one_off)` and `r max(percent_one_off)` of the
sequences with 1-nt errors were only observed once. We found that the frequency
of the most abundant 1-nt error paralleled the number of sequences.
Surprisingly, the same 1-nt error appeared
`r as.numeric(names(non_random$V1V6[length(non_random$V1V6)]))` times in the
V1-V6 mock data and another 1-nt error appeared
`r as.numeric(names(non_random$V1V9[length(non_random$V1V9)]))` times in the
V1-V9 mock data. Contrary to previous reports
{Carneiro, 2012 #3458;Koren, 2012 #3459}, these results indicate that
reproducible errors occur with the PacBio sequencing platform and that they can
be quite abundant. Through the use of the pre-clustering step described above
these 1-nt errors would be ameliorated; however, this result indicates that
caution should be used when attempting to use fine-scale OTU definitions.

## Conclusions
The various sequencing platforms that are available to microbial ecologists are
able to fill unique needs and have their own strengths and weaknesses. For
sequencing the 16S rRNA gene, the 454 platform is able to generate a moderate
number of high-quality 500-nt sequence fragments (error rates below 0.02%) {Schloss, 2011 #2466} and
the MiSeq platform is able to generate a large number of high-quality 250-nt
sequence fragments (error rates below 0.02%) {Kozich, 2013 #2719}. The promise of the PacBio
sequencing platform was the generation of high-quality near full-length sequence
fragments. As we have shown in this study, it is possible to generate near
full-length sequences; however, the error rate associated with those reads is
considerable (i.e. `r round(100*pooled_error["v19","precluster.error"],2)`%)
and requires a level of sequencing coverage that is not commonly observed in a
typical sequencing run. This results in the generation of a small number of low
quality full-length sequences. When we considered the shorter V4 region, which is
similar in length to what is sequenced by the MiSeq platform, the error rates
we observed with the PacBio platform were nearly 5-fold higher than what has
previously been reported. It appears that the promise offered by the PacBio
platform has not been realized.

@@

oligos <- read.table(file="pacbio.oligos", stringsAsFactors=F)[1:6,]
rownames(oligos) <- pretty_region[oligos$V4]
oligos <- oligos[,c("V2", "V3")]
colnames(oligos) <- c("Forward", "Reverse")

@@

```{r Table_2, warning=TRUE, results='asis'}
rownames(reasons.lost) <- pretty_region[rownames(reasons.lost)]
reasons.lost <- cbind(orig.nseqs, 100*reasons.lost, basic.nseqs)
knitr::kable(reasons.lost, digits=1, row.names=TRUE,
						 col.names=c("Initial sequences (N)", "Good reads (%)",
						 						"Wrong start/end position (%)",
						 						"Excessively long homopolymers (%)",
						 						"Ambiguous base calls (%)",
						 						"Sequences remaining (N\\)"),
						 align="c")
```

@@

par(mar=c(5,2,0.5, 0.5))
plot(1, xlim=c(2,30), ylim=c(0,0.05), type="n", xlab="", ylab="", axes=F)
axis(2, at=seq(0,0.05, 0.01), label=FALSE, tick=T, las=2)
#axis(2, at=seq(0,0.05, 0.01), label=seq(0,5, 1), las=2)
axis(1, cex.axis=1.5)
lapply(regions, plotLines)
legend(x=18, y=0.05, legend=pretty_region[regions], lty=1, lwd=2, col=clrs[regions])
mtext(side=2, line=0.5, at=0.05, text="C", cex=2, font=2, las=2)
title(xlab="Coverage", ylab="", cex.lab=1.5)
box()

@@

legend(x=10, y=0.50, legend=pretty_region[regions], lty=1, lwd=2, col=clrs[regions], cex=1.0)
```
